Current and future constraints on primordial magnetic fields

### 2017-01-30 pmf.tex
started sec 4.4.1

\begin{figure}[htb]\centering
\includegraphics[width=0.45\textwidth,clip,trim={1.5cm 12.5cm 11cm 7.5cm}]{pmf_sens.pdf}
  \caption[]{ \label{fig:ndetector}
  How the expected uncertainty on the PMF power scales with the polarized map noise. 
  Forecasts are shown for each of the three PMF scalar indices: $n_B = -2.9$ (solid), -1 (dashed), and +2 (dash-dot). 
  The black lines are for the conservative 11-parameter model, while the blue lines are for the simple \lcdm{}+\apmf{} model. 
  The model only matters in the nearly-scale invariant PMF case, where the constraints degrade in the 11-parameter model due a degeneracy with the tensor-to-scalar ratio. 
  The PMF constraints improve rapidly with lower noise levels in all three cases up to $\sim$\,1\,\ukarcmin. 
    }
\end{figure}

@@

\begin{figure}[htb]\centering
\includegraphics[width=0.45\textwidth,clip,trim={1.5cm 12.5cm 11cm 7.5cm}]{pmf_beam.pdf}
  \caption[]{ \label{fig:beam}
 Larger telescopes (and smaller beams) will improve limits on the PMF power. 
 Forecasts are shown for each of the three PMF scalar indices: $n_B = -2.9$ (solid), -1 (dashed), and +2 (dash-dot). 
  The black lines are for the conservative 11-parameter model, while the blue lines are for the simple \lcdm{}+\apmf{} model. 
  The model only matters in the nearly-scale invariant PMF case ($n_B=-2.9$), where the information on large angular scales dominates and the telescope size is relatively unimportant. 
    }
\end{figure}

Larger telescopes will allow CMB polarization to be measured on smaller angular scales, and should therefore always improve the PMF constraints. 
As shown in Fig.~\ref{fig:beam}, we find the gains due to resolution at fixed mapping speed to be substantial for all three power law indices considered.
 A beam size of FWHM=1$^\prime$ instead of 10$^\prime$ reduces the expected upper limit five-fold for $n_b=-2.9$ and 30- to 35-fold for the steeper PMF indices which are more heavily weighted towards small angular scales.

@@

In this work, we have improved the current upper limits on the strength of any primordial magnetic fields by including more CMB B-mode polarization data. 
By adding \bicepkeck{}, \pb, and \sptpol{} to \planck{} we find the 95\% CL upper limit on the PMF power falls nearly two-fold from $\apmf < \upperAplk$ to $\apmf < \upperAall$. 
The biggest contributor to this improvement is the low-$\ell$ data from \bicepkeck{}.

We have also shown that the next generation of experiments should dramatically reduce these limits, and potentially detect PMFs for the first time. 
These so-called stage III experiments, which will begin taking data in 2017, expect to set upper limits at the level of $\apmf < 0.02$ even after marginalizing over a five-parameter extension to \lcdm.

The potential for detection increases even further with planned experiments like the Simons Observatory or CMB-S4. 
We show that an ideal version of CMB-S4 might decrease the 95\% CL upper limits eight-fold, to $\apmf < 0.00025$, and even non-ideal versions can easily set upper limits on order of $\apmf < 0.006$, i.e.~ a three-fold improvement over the forecasts for the stage III experiments and a 37-fold improvement over the current limits. 
We consider how the design of these experiments will impact the resulting PMF limits -- the experimental sensitivity, angular resolution, survey area, low-frequency noise performance, and beam or calibration uncertainties -- and find sensitivity followed by telescope size to be the most important factors in predicting the PMF limits. 
CMB-S4 will be a very exciting probe of PMFs and other sources of cosmic birefringence.

>>>>

Between each stage of ground-based CMB experiment, the number of detectors increases by an order of magnitude. Adding detectors will always reduce the noise in the instrument, making it easier to isolate the PMF signal, thereby reducing the uncertainty on PMF measurements. When quantifying the effects of the number of detectors it is more informative to use polarisation noise, as detector count relies on knowing the individual sensitivity of each detector. As such, we give our result in terms of polarisation noise, noting that polarisation noise scales with detector count as $\sqrt{N_{detectors}}$

We find that

\begin{figure}[htb]\centering
\includegraphics[width=0.45\textwidth,clip,trim={1.5cm 12.5cm 11cm 7.5cm}]{pmf_sens.pdf}
  \caption[Insert plot of error on $A_{PMF}$ vs number of detectors]{ \label{fig:ndetector}
  How the expected uncertainty on the PMF power scales with the polarized map noise. 
  Forecasts are shown for each of the three PMF scalar indices: $n_B = -2.9$ (solid), -1 (dashed), and +2 (dash-dot). 
  The black lines are for the conservative 11-parameter model, while the blue lines are for the simple \lcdm{}+\apmf{} model. 
  The model only matters in the nearly-scale invariant PMF case, where the constraints degrade in the 11-parameter model due a degeneracy with the tensor-to-scalar ratio. 
  The PMF constraints improve rapidly with lower noise levels in all three cases up to $\sim$\,1\,\ukarcmin. 
    }
\end{figure}

@@

Larger telescopes will allow CMB polarization to be measured on smaller angular scales, and should therefore always improve the PMF constraints. 
We find the gains due to resolution at fixed mapping speed to be substantial for all three power law indices considered: a beam size of FWHM=1$^\prime$ instead of 10$^\prime$ reduces the expected upper limit five-fold for $n_b=-2.9$ and 30- to 35-fold for the steeper PMF indices which are more heavily weighted towards small angular scales.

@@

In this work, we have improved the current upper limits on the strength of any primordial magnetic fields by including more CMB B-mode polarization data. 
By adding \bicepkeck{}, \pb, and \sptpol{} to \planck{} we find the 95\% CL upper limit on the PMF power falls from $\apmf < \upperAplk$ to $\apmf < \upperAall$. 
The biggest contributor to this improvement is the low-$\ell$ data from \bicepkeck{}.

We have also shown that the next generation of experiments should dramatically reduce these limits, and potentially detect PMFs for the first time. 
These so-called stage III experiments, which will begin taking data in 2017, expect to set upper limits at the level of $\apmf < 0.0062$ even after marginalizing over a five-parameter extension to \lcdm.

The potential for detection increases even further with planned experiments like the Simons Observatory or CMB-S4. 
We show that an ideal version of CMB-S4 might decrease the 95\% CL upper limits eight-fold, to $\apmf < 0.0008$, and even non-ideal versions can easily set upper limits on order of $\apmf < 0.001$, i.e.~ a six-fold improvement over the forecasts for the stage III experiments and a 330-fold improvement over the current limits. 
These experiments will be very exciting probes of PMFs and other sources of cosmic birefringence.
